Malacology: A Programmable Storage System

### 2016-10-14 paper/paper.tex
paper: focus the challenges section, add transition
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{abstract}
Storage systems are caught between rapidly changing data
processing systems and the increasing speed of storage devices. This puts
tremendous pressure on storage systems to adapt both in terms of their
interfaces and their performance. But adapting storage systems can be difficult
because unprincipled changes might jeopardize years of code-hardening and
performance optimization efforts that were necessary for users to entrust their
data to the storage system.  We introduce  Malacology, a prototype programmable
storage system to explore how existing abstractions of common services found in
storage systems can be leveraged to address new data processing systems and the
increasing speed of storage devices. This approach allows a large degree of
flexibility for storage systems to evolve without sacrificing the robustness of
their code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by composing existing primitives into two new higher-level
services: a file system metadata load balancer and a high-performance distributed
shared-log that leverages flash devices. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to balance load, efficiently propogate cluster metadata, recover
from failure, and to make trade-offs for latency/throughput using leases.
\end{abstract}

@@

\item An interface for managing strongly-consistent time-varying **service
metadata**.

\item An interface for installing and evolving domain-specific, cluster-wide **data
interfaces**.

\item An interface for managing access to **shared resources** using a variety of
optimization strategies.

\item An interface for **load balancing** resources across the cluster.

\item An interface for **durability** that persists policies using the underyling storage
stack's object store.

@@

Establishing the infrastructure for programmability into existing services and
abstractions of distributed storage systems is challenging, even if one assumes
that the source code of the storage system and the necessary expertise for
understanding it is available:

@@

\item Policies and optimizations are usually hard-wired into the services and
one has to be careful when factoring them out not to introduce additional bugs.

\item Mechanisms that are only exercised according to hard-wired policies and
not in their full generality have hidden bugs that are revealed as soon as
those mechanisms are governed by different policies. In our experience
introducing programmability into a storage system proved to be a great
debugging tool.

@@

@@

Malacology is both a prototype for a programmable storage system and a design
approach to evolve storage systems efficiently and without jeopardizing years
of code-hardening and performance optimization efforts. The guiding principle
is to re-use existing services and extend them so that these services can be
\emph{programmed}. We accomplish programmability of a service by exporting
bindings (or ``hooks'') for an interpreted programming language so that
programming can occur without having to restart the storage system (see also
below,~\S\ref{sec:durability}).

>>>>

\begin{abstract}
Storage systems are caught between rapidly changing data
processing systems and the increasing speed of storage devices. This puts
tremendous pressure on storage systems to adapt both in terms of their
interfaces and their performance. But adapting storage systems can be difficult
because unprincipled changes might jeopardize years of code-hardening and
performance optimization efforts that were necessary for users to entrust their
data to the storage system.  We introduce  Malacology, a prototype programmable
storage system to explore how existing abstractions of common services found in
storage systems can be leveraged to address new data processing systems and the
increasing speed of storage devices. This approach allows a large degree of
flexibility for storage systems to evolve without sacrificing the robustness of
their code-hardened subsystems.  We illustrate the advantages and challenges of
programmability by composing existing primitives into two new higher-level
services: a file system metadata load balancer and a high-performance distributed
shared-log that leverages flash devices. The evaluation demonstrates that our
services inherit desirable qualities of the back-end storage system, including
the ability to balance load, efficiently propagate cluster metadata, recover
from failure, and to make trade-offs for latency/throughput using leases.
\end{abstract}

@@

\item An interface for managing strongly-consistent time-varying \textbf{service
metadata}.

\item An interface for installing and evolving domain-specific, cluster-wide
\textbf{data interfaces}.

\item An interface for managing access to \textbf{shared resources} using a
variety of optimization strategies.

\item An interface for \textbf{load balancing} resources across the cluster.

\item An interface for \textbf{durability} that persists policies using the
underyling storage stack's object store.

@@

Implementing the infrastructure for programmability into existing services and
abstractions of distributed storage systems is challenging, even if one assumes
that the source code of the storage system and the necessary expertise for
understanding it is available:

@@

\item Policies and optimizations are usually hard-wired into the services and
one has to be careful when factoring them out not to introduce additional bugs.
These policies and optimiations are usually cross-cutting solutions to concerns
or trade-offs that cannot be fully explored yet (as they relate to workload or
hardware). Given these policies and optimizations, decomposition of otherwise
orthogonal internal abstractions can be difficult or dangerous.

\item Mechanisms that are often only exercised according to hard-wired policies and
not in their full generality have hidden bugs that are revealed as soon as
those mechanisms are governed by different policies. In our experience
introducing programmability into a storage system proved to be a great
debugging tool.

@@

To address these challenges we present Malacology. Malacology is both a
prototype for a programmable storage system and a design approach to evolve
storage systems efficiently and without jeopardizing years of code-hardening
and performance optimization efforts.  Although Malacology uses
the internal abstractions of the underlying storage system, including its
subsystems, components, and implementations, we emphasize that our system
still addresses the general challenges outlined above.

The main challenge of designing a programmable storage system is choosing the
right internal abstractions and picking the correct layers for exposing them.
In this paper, we do not present an exhaustive list of possible internal
abstractions nor do we contend that the abstractions we choose provide the best
trade-offs for all applications. For example, if consensus is correctly
exposed, one could implement versioning, serialization, or capabilities on top;
but perhaps versioning, serialization, and capabilities would be better
interfaces for a given set of applications. These questions are not answered in
this paper and instead we focus on showing the feasibility of building such a
system, given advances in the quality and robustness of today's storage stacks.

Malacology implements and demonstrates the ideas of programmable storage in
Ceph. We choose Ceph because it is a production quality system and because it
is open source. The large developer community ensures that code is relatively
robust and the visibility of the code lets us expose any interface we want. In
the next section we describe the Ceph components that we expose as Malacology
interfaces.

@@

The guiding principle
is to re-use existing services and extend them so that these services can be
\emph{programmed}. We accomplish programmability of a service by exporting
bindings (or ``hooks'') for an interpreted programming language so that
programming can occur without having to restart the storage system (see also
below,~\S\ref{sec:durability}).
